---
title: "Application"
description: |
  Application of plotly in R.
author:
  - name: Natalie Lenberger
    url: https://nlenberger.github.io/projectblog/
date: 04-30-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
library(plotly)
library(mdsr)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

On my blog, you will find a brief tutorial for plotly. This analysis post delves deeper into the package and shows an application of the package. In the tutorial, I used the babynames data to show how to build interactive graphics. In this post, I want to use plotly and the iris data set to show more intricate and difficult graphs.

## Analysis

First, I want to play around with the mapping techniques.

```{r}
plot_ly(iris, x = ~Petal.Length, y = ~Sepal.Length, symbol = ~Species, symbols = c("circle", "x"), marker = list(size=5, opacity=0.5))
```
This graph not only maps the different species to a color, but also to different symbols. This could be beneficial for someone who is color blind. 


Now, I will play around with the hover data labels.

```{r}
plot_ly(iris, x = ~Petal.Length, y = ~Sepal.Length, text = ~paste("Sepal Length:", Sepal.Length, ",", "Petal Length:", Petal.Length), color = ~Species)
```

For some graphics, you might want the viewer to know what exact values they are looking at, for each point. 


Same thing, but let's see if we can adjust the points based off a variable

```{r}
plot_ly(iris, x = ~Petal.Length, y = ~Sepal.Length, text = ~paste("Sepal Length:", Sepal.Length, ",", "Petal Length:", Petal.Length), color = ~Species, size = ~Petal.Width)
```

This graphic allows for another variable to be shown by changing the size of the points base on the width of the petal. 

Hey! Maybe we want to add the petal width to the hover text.

```{r preview = TRUE}
plot_ly(iris, x = ~Petal.Length, y = ~Sepal.Length, text = ~paste("Sepal Length:", Sepal.Length, ",", "Petal Length:", Petal.Length, ",", "Petal Width", Petal.Width), color = ~Species, size = ~Petal.Width)
```


You can also depict this data using a histogram! 

```{r}
plot_ly(iris, x = ~Species, y = ~Sepal.Length, color = I("orange"),
        type = "box")
```


