---
title: "Application"
description: |
  Application of plotly.
author:
  - name: Natalie Lenberger
    url: https://nlenberger.github.io/projectblog/
date: 04-17-2019
output: distill::distill_article:
  self_contained: false
---

```{r setup, include=FALSE}
library(plotly)
library(mdsr)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

On my blog, you will find a brief tutorial for plotly. This analysis post delves deeper into the package and shows an application of the package. In the tutorial, I used the babynames data to show how to build interactive graphics. In this post, I want to use plotly and the iris data set to show differnt types of graphs.

## Analysis

First I want to still use a scatter plot but play around with the mapping techniques and symbols

```{r}
plot_ly(iris, x = ~Petal.Length, y = ~Sepal.Length, symbol = ~Species, symbols = c("circle", "x"), marker = list(size=5, opacity=0.5))
```

Now let's maybe look at some boxplots.

```{r}
plot_ly(iris, x = ~Species, y = ~Sepal.Length, color = I("orange"), type = "box")
```

What if I want to get more information from my boxplot?

```{r}
plot_ly(iris, x = ~Species, y = ~Sepal.Length, 
        color = ~Petal.Length, type = "box") %>% 
  layout(boxmode = "group")
```

