---
title: "Plotly"
description: |
  This article is an introduction tutorial to the package plotly.
author:
  - name: Natalie Lenberger
    url: https://nlenberger.github.io/projectblog/
date: 04-17-2019
output: distill::distill_article:
  self_contained: false
---

```{r setup, include=FALSE}
library(plotly)
library(mdsr)
library(ggplot2)
library(babynames)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This tutorial will teach you about installing and using the package plotly in R. There are a few ways in which you might create graphics using plotly:

* directly instalizing a plotly object using `plot_ly()`, `plot_geo()`, or `plot_mapbox()`

* transforming a ggplot2 graphic into a plotly object using `ggplotly()`

This tutorial will only cover directly instalizing a plotly object.

In order to adequetly teach you about this package, we will need some data. This tutorial will take you through examples using the babynames data from the packaged babynames.

## Installation

Before you get started you must first install the package plotly.

```{r}
install.packages("plotly")
```

In order to learn about

You might also want to install the package babynames in order to follow along with these examples. Or, if you have your own data you want to apply these examples to, that is fine as well.

```{r}
install.packages("babynames")
```

## Intro to `plot_ly()`

First, let's load the plotly R package

```{r}
library(plotly)
```
`
When you start to assign variable names to visual properties within `plot_ly()`, it tries to find a sensible geometric representation of the info. 

For these next few examples, we will look at just the name "Reece" from the babynames pakage and keep the graphics basic in order to unsterday what is happening. 


```{r}
babydata <- babynames %>% 
  filter(name == "Reece") %>% 
  select(name, sex, year, prop) 
```


```{r}
plot_ly(babydata, x = ~sex)
```


```{r}
plot_ly(babydata, x = ~sex, y = ~prop)
```

You can then begin to manipulate the graphic by changing the color

```{r}
plot_ly(babydata, x = ~sex, y = ~prop, color= I("red"), stroke = I("black"))
```

Without the `I()` function, you will not obtain expect/desired results. This is because the arguments map data values to a visual range by default. 

```{r}
plot_ly(babydata, x = ~sex, y = ~prop, color = "red")
```

## Modifying `plot_ly` graphics

The plotly package takes a functional approach; modification of a plotly graphic is determined by the inpute values to a function.

In the below graphic, the function `layout()` is anticipating a plotly object for the first argument. The following arguments add or modfy the layout components. 

```{r, knitr::opts_chunk$set(echo = FALSE}
layout(
  plot_ly(babydata, x = ~sex),
  title = "Histogram for the name Reece"
)
```

This could start to get out of hand when you start wanting to make numerous modifications. For that reason, the pipe operator, `%>%`, should be used when making modifcations. This will also make it easier for other people to follow. 

```{r}
babydata %>% 
  plot_ly(x = ~sex) %>% 
  layout(title = "Histogram for the name Reece")
```

There are several other modification functions that can be used in the plotly package in order to add a graphical layer to a plot. 

A few of these include:
 * `add_histogram()`
 * `add_bars()`
 * `add_lines()`
 * `add_text()`
 * `add_contour()`
 * `add_boxplot()`
 * `add_markers()
 * `add_paths()`
 * `add_segments()`
 * `add_ribbons()`
 * `add_area()`
 * `add_polygons()`
....and many more!
 
For most graphics you create, you will want to combain several of these graphical layers. When doing that, keep in mind that any argument specified in `plot_ly()` is global. This means that any additional function will inherit that argument. If you would like to change this, you must set `inherit = FALSE`. 


Let's try adding some modifications to your simple examples.

```{r}
babydata %>%
  plot_ly(x = ~sex) %>%
  add_histogram() %>%
  group_by(sex) %>%
  summarise(n = n()) %>%
  add_text(
    text = ~scales::comma(n), y = ~n,
    textposition = "top middle",
    cliponaxis = FALSE
  )
```


## Some more complex graphics

When examining the babynames dataset, using a histogram might not be the best way to represent a greater portion of the data. 

Let's try some scatter plots!

```{r}
pal <- c("hot pink", "blue")

plot_ly(babydata, x = ~year, y = ~prop, color = ~sex, colors = pal, marker = list(size=5, opacity = 0.5)) %>% 
  layout(title= "Proportion of baby's with the name Reece by year")
```

This graph shows the proportion of baby's with the name Reece over the years. The color parameter alows for you to separate variables by color; in this case, it separates the babies by sex. You can also chose the colors you wish to depict them as using the colors parameter. The marker paramter alows you to adjust of the points on the graph. You can adjust the size, color and opacity as shown above. The layout function is how we are able to attach a title to the graph. You can see here that you also have the ability to hover over points in order to get the exact numbner.

## You try

Make an interactive graph similar to the one above using your own name. Make sure the viewer can see the difference in sex. 





